SMT-Based Context-Bounded Model Checking for Embedded Systems: Challenges and Future Trends

### 2016-03-27 paper.tex
corrections to the entire text

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The dependency on the correct functioning of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control ({\it e.g.}, airbag control), health care, surveillance, mobile devices, and consumer electronics. Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, in order to meet increasing computational power demands. As a consequence, reliability of embedded (distributed) software becomes a key issue during system development, which must be carefully addressed and assured. Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation; however, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, and supporting legacy designs (usually written in low-level languages). The present paper discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems. Reliability issues, in the development of micro-grids and cyber-physical systems, are then considered, as a prominent (bounded) model checking application.
\end{abstract}

@@

One promising approach to achieve completeness, in BMC techniques, is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. The main challenge, however, of such an approach relies on computing and strengthening inductive invariants from programs. In particular, loop invariants, which were computed from programs under verifications, must be inductive (and not just invariant) for the corresponding verification conditions to be valid, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}. As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that verifiers can automatically prove program correctness.

@@

However, there is still little evidence, in the available literature, that model checking hardware and software systems, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded-system verification. That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes. Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants about memory and variable aliasing~\cite{Bradley07}.

@@

This position paper tackles four major problems in computer-aided verification for embedded systems, which are (partially) open in current published research.

@@

\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in the literature, which demonstrate its effectiveness in some real-life embedded-systems applications~\cite{Gadelha15,Brain15,Rocha15,Donaldson10}; however, the main challenge still remains open, {\it i.e.}, to compute and strengthen loop invariants to prove program correctness and timeliness in a more efficient and effective way, in order to be competitive with other model-checking approaches. In particular, invariant-generation algorithms have substantially evolved over the last years to discover inductive invariants of programs~\cite{pips:2013,Henry:2012} or to continuously refine them during verification~\cite{Beyer15}. However, there is still a lack of studies for exploiting the combination of different invariant-generation algorithms ({\it e.g.}, interval analysis, linear inequalities, polynomial equalities and inequalities) and how to strengthen them during verification, in order to ensure system robustness w.r.t. implementation aspects.

\textbf{(RP4)} State-of-the-art SMT-based context-BMC approaches were extended to verify overflow, limit cycle, time constraints, stability, and minimum phase, in digital systems. Indeed, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter} were tackled, in order to specify system-level properties of those systems using linear-time temporal logic~\cite{JMorse15}. In particular, a specific UAV application was tackled, with the goal to verify its attitude controllers~\cite{Bessa16}. In general, however, there is still a lack of studies to verify system-level properties related to embedded systems; emphasis should be given to micro-grids~\cite{xu15} and cyber-physical systems~\cite{leeCPS2}, which require high-dependability requirements for computation, control, and communication. Additionally, the application of automated fault detection, localization, and correction techniques to digital systems represents an important research direction to make BMC tools useful for embedded systems engineers~\cite{Alves15}.

\textcolor{blue}{\textbf{(RP5)} Although ESMBMC \cite{Cordeiro12} was extended to support C/C++ and some variants, new application interfaces and languages are often developed, which then claim for suitable verification tools. Indeed, it would be interesting if one needed only to load a new language model, which along with a BMC core could check new programs. Some work towards that was already presented by Sousa, Cordeiro, and Filho, which employed operational models  for checking Qt-based programs. In summary, the BMC core (in that case, ESBMC) is not changed, but instead an operation model, which implements the behavior and features of Qt libraries, is used for providing the code structure to be checked. Such research problem is closely related to the first one (\textbf{(RP1)}) and has the potential to devise a new paradigm in software verification.}

Lastly, yet importantly, BMC tools like CBMC~\cite{Clarke04}, ESBMC~\cite{MorseCNF13,MorseRCN014}, and LLBMC~\cite{MerzFS12} represent the most prominent ones for verifying C programs, as observed in the Intl. Competition on Software Verification~\cite{Beyer14,BeyerSVCOMP15}, where verifiable (correctness and violation) witnesses are of extreme importance for evaluating software verifiers~\cite{BeyerW15,RochaIFM12}.

@@

This paper presented the main challenges related to the verification of design correctness in embedded systems, and also raises some important side considerations. In particular, it emphasizes that stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency), along with system behavior models, must be considered during verification. Additionally, there is little evidence that model checking embedded software using \textit{k}-induction (and invariants), which
extends BMC-based approaches from falsification to verification, can be applied to formally verify correctness and timeliness of embedded systems.

Given that software complexity has significantly increased in embedded products, there are still some (recent) advances to stress and exhaustively cover the system state space, in order to verify low-level properties that have to meet the application's deadline, access memory regions, handle concurrency, and control hardware registers. Besides, there is a trend towards incorporating knowledge about the system to be verified, which may take software verification one step further, where not only code correctness will be addressed, but also full system reliability.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The dependency on the correct functioning of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control ({\it e.g.}, airbag control), health care, surveillance, mobile devices, and consumer electronics. Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory, in order to meet increasing computational power demands. As a consequence, reliability of embedded (distributed) software becomes a key issue during system development, which must be carefully addressed and assured. Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation; however, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, and supporting new software architectures and legacy designs (usually written in low-level languages). The present paper discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems. Reliability issues, in the development of micro-grids and cyber-physical systems, are then considered, as a prominent (bounded) model checking application.
\end{abstract}

@@

One promising approach to achieve completeness, in BMC techniques, is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. The main challenge, however, of such an approach relies on computing and strengthening inductive invariants from programs. In particular, loop invariants, which were computed from programs under verification, must be inductive (and not just invariant) for the corresponding verification conditions to be valid, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}. As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that verifiers can automatically prove program correctness.

@@

Nonetheless, there is still little evidence, in the available literature, that model checking hardware and software systems, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded-system verification. That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes. Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants about memory and variable aliasing~\cite{Bradley07}.

@@

This position paper tackles five major problems in computer-aided verification for embedded systems, which are (partially) open in current published research.

@@

\textbf{(RP3)} Novel approaches to model check embedded software using \textit{k}-induction and invariants were proposed and evaluated in the literature, which demonstrate its effectiveness in some real-life embedded-system applications~\cite{Gadelha15,Brain15,Rocha15,Donaldson10}; however, the main challenge still remains open, {\it i.e.}, to compute and strengthen loop invariants to prove program correctness and timeliness in a more efficient and effective way, in order to be competitive with other model-checking approaches. In particular, invariant-generation algorithms have substantially evolved over the last years, with the goal of discovering inductive invariants of programs~\cite{pips:2013,Henry:2012} or continuously refine them during verification~\cite{Beyer15}; however, there is still a lack of studies for exploiting the combination of different invariant-generation algorithms ({\it e.g.}, interval analysis, linear inequalities, polynomial equalities and inequalities) and how to strengthen them during verification, in order to ensure system robustness w.r.t. implementation aspects.

\textbf{(RP4)} State-of-the-art SMT-based context-BMC approaches were extended to verify overflow, limit cycle, time constraints, stability, and minimum phase, in digital systems. Indeed, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter} were tackled, in order to specify system-level properties of those systems, using linear-time temporal logic~\cite{JMorse15}. In particular, a specific UAV application was tackled, with the goal to verify its attitude controllers~\cite{Bessa16}. In general, however, there is still a lack of studies to verify system-level properties related to embedded systems; emphasis should be given to micro-grids~\cite{xu15} and cyber-physical systems~\cite{leeCPS2}, which require high-dependability requirements for computation, control, and communication. Additionally, the application of automated fault detection, localization, and correction techniques to digital systems represents an important research direction to make BMC tools useful for embedded systems engineers~\cite{Alves15}.

\textcolor{blue}{\textbf{(RP5)} Although ESMBMC \cite{Cordeiro12} was extended to support C/C++ and some variants, new application interfaces and languages are often developed, which then claim for suitable verification tools. Indeed, it would be interesting if one needed only to load a new language model, which along with a BMC core could check different programs. Some work towards that was already presented by Sousa, Cordeiro, and Filho, which employed operational models for checking Qt-based programs. In summary, the BMC core (in that case, ESBMC) is not changed, but instead an operation model, which implements the behavior and features of Qt libraries, is used for providing the new code structure to be checked. Such research problem is closely related to the first one \textbf{(RP1)} and has the potential to devise a new paradigm in software verification.}

Lastly, yet importantly, BMC tools like CBMC~\cite{Clarke04}, ESBMC~\cite{MorseCNF13,MorseRCN014}, and LLBMC~\cite{MerzFS12} represent the most prominent approaches for verifying C programs, as observed in the Intl. Competition on Software Verification~\cite{Beyer14,BeyerSVCOMP15}, where verifiable (correctness and violation) witnesses are of extreme importance for evaluating software verifiers~\cite{BeyerW15,RochaIFM12}.

@@

This paper presented the main challenges related to the verification of design correctness, in embedded systems, and also raised some important side considerations. In particular, it emphasizes that stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency), along with system behavior models, must be considered during verification. Additionally, there is little evidence that model checking embedded software using \textit{k}-induction (and invariants), which
extends BMC-based approaches from falsification to verification, can be applied to formally verify correctness and timeliness of embedded systems.

Given that software complexity has significantly increased in embedded products, there are still some (recent) advances to stress and exhaustively cover the system state space, in order to verify low-level properties that have to meet the application's deadline, access memory regions, handle concurrency, and control hardware registers. Besides, there is a trend towards incorporating knowledge about the system to be verified, which may take software verification one step further, where not only code correctness will be addressed, but also full system reliability. \textcolor{blue}{Finally, it seems interesting to provide behavioral models when new application interfaces or language features are used, in order to extend the capabilities of current verification tools, without changing the core BMC module.}
